---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path  = "man/figures/README-",
  out.width = "100%",
  eval      = FALSE
)
```

# ontologyDiscoverR

**AI-assisted ontology discovery from documents and schemas.**

`ontologyDiscoverR` uses the Anthropic Claude API to automatically extract
ontology structure — object types, properties, link types, action types, and
concept definitions — from database schemas, API specs, documentation, and
data files. The output is a validated bundle ready for use with `ontologySpecR`,
`objectSetsR`, and `conceptR`.

## Installation

```{r install}
# install.packages("pak")
pak::pkg_install("cathalbyrnegit/ontologydiscoverr")
```

Set your Anthropic API key before use:

```{r apikey}
Sys.setenv(ANTHROPIC_API_KEY = "your-key-here")
# Or add ANTHROPIC_API_KEY=your-key-here to your ~/.Renviron
```

## How it works

```
Source documents / schemas / APIs
          ↓
     ontologyDiscoverR
       ├── parse      (extract text/structure per source type)
       ├── extract    (LLM → candidate types, properties, links)
       ├── merge      (deduplicate, resolve conflicts)
       ├── review     (human approval Shiny app)
       └── emit       (ontologySpecR bundle)
          ↓
     ontologySpecR bundle → rest of stack
```

The LLM runs **three sequential passes** per source:

1. **Object types** — entities with identity (tables, schema components, named domain concepts)
2. **Link types** — relationships between object types (FK constraints, API references, document links)
3. **Action types & concept hints** — verbs/endpoints and business-rule conditions

## Quick start

```{r quickstart}
library(ontologyDiscoverR)

# 1. Create a session and add sources
sess <- discover_session("hospital-demo")
sess <- dis_add_file(sess, "path/to/hospital-schema.sql")
sess <- dis_add_file(sess, "path/to/api-spec.yaml")
sess <- dis_add_file(sess, "path/to/domain-docs.pdf")

# 2. Run extraction (calls Claude API)
sess <- dis_extract(sess, verbose = TRUE)
#> ℹ Processing source 1/3: hospital-schema.sql
#> ℹ Processing source 2/3: api-spec.yaml
#> ℹ Processing source 3/3: domain-docs.pdf
#> ℹ Merging and deduplicating candidates...
#> ℹ Sources processed: 3
#> ℹ Object types found: 12 (8 high-confidence, 4 medium, 0 low)
#> ℹ Link types found: 7
#> ℹ Action types found: 4
#> ℹ Concept hints found: 6
#> ℹ Conflicts requiring review: 2

# 3. Review candidates interactively (optional)
sess <- dis_review(sess)

# 4. Export to an ontologySpecR bundle
bundle <- dis_to_bundle(
  sess,
  bundle_id   = "hospital-v1",
  bundle_name = "Hospital Management Ontology"
)
```

## Supported source types

| Source type | Function | Notes |
|---|---|---|
| SQL DDL file | `parse_sql_ddl()` | CREATE TABLE → object types, FK → link types |
| Live database | `parse_db_schema()` | DBI connection, reads information_schema |
| OpenAPI spec | `parse_openapi()` | 3.x + Swagger 2.x, YAML or JSON |
| CSV file | `parse_csv()` | Header + sample rows, type inference |
| R data frame | `parse_r_dataframe()` | In-memory, includes `str()` for LLM context |
| PDF | `parse_pdf()` | Text extracted via `pdftools`, up to 50k chars |
| Markdown / HTML | `parse_markdown()` / `parse_html()` | Heading structure preserved |
| OWL / RDF | `parse_owl()` | Classes → object types, properties → links |

## Live database example

```{r db}
library(DBI)
library(duckdb)

con  <- dbConnect(duckdb::duckdb(), "hospital.ddb")
sess <- discover_session("from-db")
sess <- dis_add_db(sess, con, label = "hospital-db")
sess <- dis_extract(sess)
```

## Confidence scores

Every extracted candidate carries a confidence score (0–1):

| Score | Meaning |
|---|---|
| 1.0 | Explicitly defined in schema / data dictionary |
| 0.9 | Clearly named in documentation with attributes listed |
| 0.7 | Mentioned as a concept with some structure implied |
| 0.5 | Inferred from context |
| 0.2 | Speculative |

Use `min_confidence` in `dis_to_bundle()` to control which pending candidates
are included:

```{r confidence}
bundle <- dis_to_bundle(sess, "my-bundle", "My Ontology", min_confidence = 0.7)
```

## Interactive review

```{r review}
sess <- dis_review(sess)
```

The Shiny app has four panels:

- **Object Types** — approve/reject individual candidates or bulk-approve all ≥ 0.8 confidence
- **Link Types** — correct from/to endpoints if the LLM misidentified them
- **Concept Hints** — edit the suggested SQL expression inline, test against a live DB
- **Conflicts** — resolve pairs of conflicting candidates side by side

## MCP conversational building

If `ontologyMCP` is installed, you can build the ontology interactively:

```{r mcp}
library(ontologyMCP)

server <- mcp_server("ontology-builder")
server <- dis_add_mcp_tool(server, sess)
mcp_start(server)

# Now an LLM agent can say things like:
# "Add a link from Hospital to Patient called Admission"
# "The Patient object type should have a concept called high_risk where age > 65"
# "Looks good — generate the bundle"
```

## Testing

All tests mock `call_claude()` — no API key required, no cost:

```{r test}
# install.packages("mockery")
devtools::test()
```

## Package structure

```
R/
  session.R           # discover_session, dis_add_*, dis_extract, dis_summary, dis_to_bundle
  llm.R               # call_claude (Anthropic API via httr2)
  parsers/            # one file per source type
  extract.R           # four LLM extraction passes
  merge.R             # merge_candidates, detect_conflicts
  candidates.R        # S3 constructors for all four candidate types
  review.R            # dis_review() Shiny app
  mcp_builder.R       # dis_add_mcp_tool() (optional)
  types.R             # normalise_type, validate_candidate
  utils.R             # UUID, truncation, JSON helpers

inst/fixtures/        # example-schema.sql, example-openapi.yaml,
                      # example-docs.md, example-data.csv

tests/testthat/       # mocked-LLM tests for all modules
vignettes/            # four how-to vignettes
```

## License

MIT
