---
title: "Extracting an Ontology from PDF Documentation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting an Ontology from PDF Documentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## When you only have documents

Sometimes you don't have a database schema — only PDF manuals, Word documents
converted to PDF, or markdown specifications. `ontologyDiscoverR` handles this
via the `parse_pdf()` and `parse_markdown()` parsers.

```{r setup}
library(ontologyDiscoverR)
Sys.setenv(ANTHROPIC_API_KEY = "your-key-here")
```

## Adding PDF and Markdown sources

```{r sources}
sess <- discover_session("docs-demo")

# Add a PDF manual
sess <- dis_add_file(sess, "hospital-manual.pdf")

# Add a Markdown API specification
sess <- dis_add_file(sess, "api-spec.md")

# Mix with a schema if available
sess <- dis_add_file(sess, "schema.sql")
```

## Extraction and review

```{r extract}
sess <- dis_extract(sess, verbose = TRUE)
dis_summary(sess)
```

Because documentation sources are less structured than schemas, confidence
scores will typically be lower (0.5–0.8). Use `dis_review()` to validate the
results before building the bundle.

```{r review}
sess   <- dis_review(sess)
bundle <- dis_to_bundle(sess, "docs-bundle", "Documentation Ontology",
                         min_confidence = 0.6)
```

## Tips for better extraction from documents

- Include a **glossary** or **data dictionary** section if one exists — these
  give the LLM high-confidence anchors.
- For large PDFs (> 50 pages), split by chapter and run separate sessions,
  then merge candidates manually.
- Use `include_concepts = TRUE` in `dis_to_bundle()` if the documentation
  describes business rules you want to capture as concept definitions.
