---
title: "From Database Schema to Ontology Bundle in 5 Minutes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{From Database Schema to Ontology Bundle in 5 Minutes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

## Overview

This vignette demonstrates how to extract an ontology bundle from a SQL DDL
file in a few lines of R code.

## Prerequisites

```{r setup}
library(ontologyDiscoverR)

# Ensure your API key is set
Sys.setenv(ANTHROPIC_API_KEY = "your-key-here")  # or use .Renviron
```

## Step 1: Create a session and add your schema

```{r session}
sess <- discover_session("hospital-demo")
sess <- dis_add_file(sess, "path/to/hospital-schema.sql")
```

You can also point at a live database:

```{r db}
library(DBI)
library(duckdb)

con  <- dbConnect(duckdb::duckdb(), "hospital.ddb")
sess <- dis_add_db(sess, con, label = "hospital-db")
```

## Step 2: Run extraction

```{r extract}
sess <- dis_extract(sess, verbose = TRUE)
```

This runs three LLM passes per source:
1. **Pass 1**: Extract object types and their properties
2. **Pass 2**: Extract link types (relationships between objects)
3. **Pass 3**: Extract action types and concept hints

After extraction, candidates are deduplicated and merged across sources.

## Step 3: Review (optional)

```{r review, eval=FALSE}
sess <- dis_review(sess)
```

The Shiny review app lets you approve or reject each candidate, edit properties,
and resolve conflicts before building the bundle.

## Step 4: Export to bundle

```{r bundle}
bundle <- dis_to_bundle(
  sess,
  bundle_id   = "hospital-v1",
  bundle_name = "Hospital Management Ontology"
)

# The bundle is a plain list — write it to JSON for use with ontologySpecR
jsonlite::write_json(bundle, "hospital-bundle.json", auto_unbox = TRUE, pretty = TRUE)
```

## Quick summary

```{r summary}
dis_summary(sess)
```
