---
title: "Interactive Review with dis_review()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interactive Review with dis_review()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## The review workflow

After `dis_extract()` produces candidate ontology elements, human review
ensures quality before the bundle is built. The `dis_review()` Shiny app
provides a structured interface for this.

```{r setup}
library(ontologyDiscoverR)
```

## Launching the reviewer

```{r review}
# Assumes sess has been populated via dis_extract()
sess <- dis_review(sess)
```

The app opens in your browser with four panels:

### Panel 1: Object Types

Each row shows a candidate object type with its confidence score and source
count. Click a row to see:
- All properties with their types and nullability
- Source excerpts that led to this candidate
- Any conflicts with other candidates

Bulk actions:
- **Approve all ≥ 0.8**: Approves high-confidence candidates in one click
- **Reject all < 0.3**: Clears out speculative entries

### Panel 2: Link Types

Shows `from → to` relationships. The confidence-based colouring helps you
quickly spot uncertain links. Click a row to correct the from/to endpoints
if the LLM misidentified them.

### Panel 3: Concept Hints

Each hint shows the LLM's suggested SQL expression. You can edit the SQL
inline and click **Test** (if a database connection is active) to see the
TRUE/FALSE distribution. Approve hints to include them as `concept_defs` in
the bundle.

### Panel 4: Conflicts

Pairs of conflicting candidates appear side by side. For each pair, choose:
- **Keep A** — discard B
- **Keep B** — discard A
- **Keep Both** — retain both as separate candidates

## Closing the app

Click the **Done** button in the navbar. The app closes and `dis_review()`
returns the updated session with all your status changes applied.

```{r bundle}
bundle <- dis_to_bundle(sess, "reviewed-bundle", "Reviewed Ontology")
```

## Programmatic approval

If you want to auto-approve without opening the UI:

```{r programmatic}
sess$candidates$object_types <- lapply(
  sess$candidates$object_types,
  function(ot) {
    if (ot$confidence >= 0.8) ot$status <- "approved"
    ot
  }
)
```
