---
title: "Building an Ontology via LLM Conversation (MCP)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Building an Ontology via LLM Conversation (MCP)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## Iterative ontology construction via conversation

When `ontologyMCP` is installed, you can expose the session as an MCP tool
and build the ontology interactively through a conversation with Claude.

```{r setup}
library(ontologyDiscoverR)
# install.packages("ontologyMCP")  # optional companion package
library(ontologyMCP)
Sys.setenv(ANTHROPIC_API_KEY = "your-key-here")
```

## Setting up the MCP server

```{r mcp}
sess   <- discover_session("mcp-demo")
server <- ontologyMCP::mcp_server("ontology-builder")
server <- dis_add_mcp_tool(server, sess)
ontologyMCP::mcp_start(server)
```

## Example conversation flow

Once the server is running, an LLM agent can interact with it:

```
User: "Look at this database schema and build me an ontology for a hospital"
Claude: [parses schema, calls dis_extract, lists discovered types]

User: "The Patient table also connects to the Ward table via ward_assignments"
Claude: [calls build_ontology("Add a link from Patient to Ward called ward_assignments, many_to_many")]
→ Done: Added link type 'ward_assignments'

User: "Add a concept for high-risk patients: age > 65 OR previous_admissions > 0"
Claude: [calls build_ontology("Add concept high_risk_patient for Patient where age > 65 OR previous_admissions > 0")]
→ Done: Added concept hint 'high_risk_patient'

User: "Looks good — approve everything and generate the bundle"
Claude: [calls build_ontology("Approve all pending object types"), then dis_to_bundle()]
```

## Supported natural language instructions

The `build_ontology` tool understands:

| Instruction pattern | Effect |
|---|---|
| `"Add an object type for X with properties a, b, c"` | Adds a CandidateObjectType |
| `"Add a link from X to Y called Z"` | Adds a CandidateLinkType |
| `"Add a concept called X for Y where <SQL condition>"` | Adds a CandidateConceptHint |
| `"Add an action called X on Y"` | Adds a CandidateActionType |
| `"Approve all pending object types"` | Sets status = approved |
| `"Reject the concept high_risk"` | Sets status = rejected for that id |

## Retrieving the final session

After the conversation, retrieve the updated session and build the bundle:

```{r final}
final_sess <- ontologyMCP::mcp_get_session(server, "ontology-builder")
bundle     <- dis_to_bundle(final_sess, "mcp-bundle", "MCP-built Ontology",
                             include_concepts = TRUE)
```
